---
title: "PLSC 504 Hemauer Assignment 5"
output:
  pdf_document: default
  html_notebook: default
---

Part 1 

1.

```{r}

library(plm)
library(tidyr)

set.seed(1337)

sim <- function(){
  
  n = 10
  t = 10
  a = rep(rnorm(n), each = t)
  x <- rnorm(n*t)
  u <- rnorm(100)
  B0 <- 2
  B1 <- 5

  Y <- B0 + B1 * x + a + u

  ID <- rep(1:n, each = t)
  pandata <- cbind(Y, x, ID)
  pandata <- pdata.frame(pandata, index = "ID")

  within <- plm(Y ~ x, data = pandata, model = "within")
  random <- plm(Y ~ x, data = pandata, model = "random")

  withCoef <- within$coefficients
  randCoef <- random$coefficients[2]
  
  return(withCoef)

}

data <- replicate(100, sim())

sim <- function(){
  
  n = 10
  t = 10
  a = rep(rnorm(n), each = t)
  x <- rnorm(n*t)
  u <- rnorm(100)
  B0 <- 2
  B1 <- 5

  Y <- B0 + B1 * x + a + u

  ID <- rep(1:n, each = t)
  pandata <- cbind(Y, x, ID)
  pandata <- pdata.frame(pandata, index = "ID")

  within <- plm(Y ~ x, data = pandata, model = "within")
  random <- plm(Y ~ x, data = pandata, model = "random")

  withCoef <- within$coefficients
  randCoef <- random$coefficients[2]
  
  return(randCoef)

}

data2 <- replicate(100, sim())

fullData <- cbind(data, data2)

```

The difference between my fixed effects and my random effects models is minimal. The X coefficient estimates are the same, along with the standard errors.

C.

```{r}

sim <- function(){
  
  n = 10
  t = 10
  a = rep(rnorm(n), each = t)
  x <- rnorm(n*t)
  u <- rnorm(100)
  B0 <- 2
  B1 <- 5

  Y <- B0 + B1 * x + a + u

  ID <- rep(1:n, each = t)
  pandata <- cbind(Y, x, ID)
  pandata <- pdata.frame(pandata, index = "ID")

  within <- plm(Y ~ x, data = pandata, model = "within")
  random <- plm(Y ~ x, data = pandata, model = "random")

  withCoef <- within$coefficients
  randCoef <- random$coefficients[2]
  
  test <- phtest(within, random)
  
  return(test)

}

test <- replicate(100, sim())

```

The Hausman test reports in every case random effect. Which makes sense given the data generation process.

2.

```{r}

sim <- function(){
  
  n = 100
  t = 10
  a = rep(rnorm(n), each = t)
  x <- rnorm(n*t)
  u <- rnorm(100)
  B0 <- 2
  B1 <- 5

  Y <- B0 + B1 * x + a + u

  ID <- rep(1:n, each = t)
  pandata <- cbind(Y, x, ID)
  pandata <- pdata.frame(pandata, index = "ID")

  within <- plm(Y ~ x, data = pandata, model = "within")
  random <- plm(Y ~ x, data = pandata, model = "random")

  withCoef <- within$coefficients
  randCoef <- random$coefficients[2]
  
  return(withCoef)

}

data <- replicate(100, sim())

sim <- function(){
  
  n = 100
  t = 10
  a = rep(rnorm(n), each = t)
  x <- rnorm(n*t)
  u <- rnorm(100)
  B0 <- 2
  B1 <- 5

  Y <- B0 + B1 * x + a + u

  ID <- rep(1:n, each = t)
  pandata <- cbind(Y, x, ID)
  pandata <- pdata.frame(pandata, index = "ID")

  within <- plm(Y ~ x, data = pandata, model = "within")
  random <- plm(Y ~ x, data = pandata, model = "random")

  withCoef <- within$coefficients
  randCoef <- random$coefficients[2]
  
  return(randCoef)

}

data2 <- replicate(100, sim())

fullData <- cbind(data, data2)

sim <- function(){
  
  n = 100
  t = 10
  a = rep(rnorm(n), each = t)
  x <- rnorm(n*t)
  u <- rnorm(100)
  B0 <- 2
  B1 <- 5

  Y <- B0 + B1 * x + a + u

  ID <- rep(1:n, each = t)
  pandata <- cbind(Y, x, ID)
  pandata <- pdata.frame(pandata, index = "ID")

  within <- plm(Y ~ x, data = pandata, model = "within")
  random <- plm(Y ~ x, data = pandata, model = "random")

  withCoef <- within$coefficients
  randCoef <- random$coefficients[2]
  
  test <- phtest(within, random)
  
  return(test)

}

test <- replicate(100, sim())

##############################

sim <- function(){
  
  n = 10
  t = 109
  a = rep(rnorm(n), each = t)
  x <- rnorm(n*t)
  u <- rnorm(100)
  B0 <- 2
  B1 <- 5

  Y <- B0 + B1 * x + a + u

  ID <- rep(1:n, each = t)
  pandata <- cbind(Y, x, ID)
  pandata <- pdata.frame(pandata, index = "ID")

  within <- plm(Y ~ x, data = pandata, model = "within")
  random <- plm(Y ~ x, data = pandata, model = "random")

  withCoef <- within$coefficients
  randCoef <- random$coefficients[2]
  
  return(withCoef)

}

data <- replicate(100, sim())

sim <- function(){
  
  n = 10
  t = 100
  a = rep(rnorm(n), each = t)
  x <- rnorm(n*t)
  u <- rnorm(100)
  B0 <- 2
  B1 <- 5

  Y <- B0 + B1 * x + a + u

  ID <- rep(1:n, each = t)
  pandata <- cbind(Y, x, ID)
  pandata <- pdata.frame(pandata, index = "ID")

  within <- plm(Y ~ x, data = pandata, model = "within")
  random <- plm(Y ~ x, data = pandata, model = "random")

  withCoef <- within$coefficients
  randCoef <- random$coefficients[2]
  
  return(randCoef)

}

data2 <- replicate(100, sim())

fullData <- cbind(data, data2)

sim <- function(){
  
  n = 10
  t = 100
  a = rep(rnorm(n), each = t)
  x <- rnorm(n*t)
  u <- rnorm(100)
  B0 <- 2
  B1 <- 5

  Y <- B0 + B1 * x + a + u

  ID <- rep(1:n, each = t)
  pandata <- cbind(Y, x, ID)
  pandata <- pdata.frame(pandata, index = "ID")

  within <- plm(Y ~ x, data = pandata, model = "within")
  random <- plm(Y ~ x, data = pandata, model = "random")

  withCoef <- within$coefficients
  randCoef <- random$coefficients[2]
  
  test <- phtest(within, random)
  
  return(test)

}

test <- replicate(100, sim())

######################################

sim <- function(){
  
  n = 100
  t = 100
  a = rep(rnorm(n), each = t)
  x <- rnorm(n*t)
  u <- rnorm(100)
  B0 <- 2
  B1 <- 5

  Y <- B0 + B1 * x + a + u

  ID <- rep(1:n, each = t)
  pandata <- cbind(Y, x, ID)
  pandata <- pdata.frame(pandata, index = "ID")

  within <- plm(Y ~ x, data = pandata, model = "within")
  random <- plm(Y ~ x, data = pandata, model = "random")

  withCoef <- within$coefficients
  randCoef <- random$coefficients[2]
  
  return(withCoef)

}

data <- replicate(100, sim())

sim <- function(){
  
  n = 100
  t = 100
  a = rep(rnorm(n), each = t)
  x <- rnorm(n*t)
  u <- rnorm(100)
  B0 <- 2
  B1 <- 5

  Y <- B0 + B1 * x + a + u

  ID <- rep(1:n, each = t)
  pandata <- cbind(Y, x, ID)
  pandata <- pdata.frame(pandata, index = "ID")

  within <- plm(Y ~ x, data = pandata, model = "within")
  random <- plm(Y ~ x, data = pandata, model = "random")

  withCoef <- within$coefficients
  randCoef <- random$coefficients[2]
  
  return(randCoef)

}

data2 <- replicate(100, sim())

fullData <- cbind(data, data2)

sim <- function(){
  
  n = 100
  t = 100
  a = rep(rnorm(n), each = t)
  x <- rnorm(n*t)
  u <- rnorm(100)
  B0 <- 2
  B1 <- 5

  Y <- B0 + B1 * x + a + u

  ID <- rep(1:n, each = t)
  pandata <- cbind(Y, x, ID)
  pandata <- pdata.frame(pandata, index = "ID")

  within <- plm(Y ~ x, data = pandata, model = "within")
  random <- plm(Y ~ x, data = pandata, model = "random")

  withCoef <- within$coefficients
  randCoef <- random$coefficients[2]
  
  test <- phtest(within, random)
  
  return(test)

}

test <- replicate(100, sim())

```

When N is increased from 10 to 100, nothing changes from the previous steps. When T is pushed from 10 to 100, and N remains at 10, nothing is changed. When N and T are pushed to 100, the estimates are much closer together and the standard errors are closer to 0, but the Hausman test does not change.

3.

```{r}

library(plm)
set.seed(1337)

sim <- function(){
  
  n = 10
  t = 10
  a = rep(rnorm(n), each = t)
  x <- (a + rnorm(n*t)) / 2
  u <- rnorm(100)
  B0 <- 2
  B1 <- 5

  Y <- B0 + B1 * x + a + u

  ID <- rep(1:n, each = t)
  pandata <- cbind(Y, x, ID)
  pandata <- pdata.frame(pandata, index = "ID")

  within <- plm(Y ~ x, data = pandata, model = "within")
  random <- plm(Y ~ x, data = pandata, model = "random")

  withCoef <- within$coefficients
  randCoef <- random$coefficients[2]
  
  return(withCoef)

}

data <- replicate(100, sim())

sim <- function(){
  
  n = 10
  t = 10
  a = rep(rnorm(n), each = t)
  x <- (a + rnorm(n*t)) / 2
  u <- rnorm(100)
  B0 <- 2
  B1 <- 5

  Y <- B0 + B1 * x + a + u

  ID <- rep(1:n, each = t)
  pandata <- cbind(Y, x, ID)
  pandata <- pdata.frame(pandata, index = "ID")

  within <- plm(Y ~ x, data = pandata, model = "within")
  random <- plm(Y ~ x, data = pandata, model = "random")

  withCoef <- within$coefficients
  randCoef <- random$coefficients[2]
  
  return(randCoef)

}

data2 <- replicate(100, sim())

fullData <- cbind(data, data2)

######################################

sim <- function(){
  
  n = 100
  t = 100
  a = rep(rnorm(n), each = t)
  x <- (a + rnorm(n*t)) / 2
  u <- rnorm(100)
  B0 <- 2
  B1 <- 5

  Y <- B0 + B1 * x + a + u

  ID <- rep(1:n, each = t)
  pandata <- cbind(Y, x, ID)
  pandata <- pdata.frame(pandata, index = "ID")

  within <- plm(Y ~ x, data = pandata, model = "within")
  random <- plm(Y ~ x, data = pandata, model = "random")

  withCoef <- within$coefficients
  randCoef <- random$coefficients[2]
  
  return(withCoef)

}

data <- replicate(100, sim())

sim <- function(){
  
  n = 100
  t = 100
  a = rep(rnorm(n), each = t)
  x <- (a + rnorm(n*t)) / 2
  u <- rnorm(100)
  B0 <- 2
  B1 <- 5

  Y <- B0 + B1 * x + a + u

  ID <- rep(1:n, each = t)
  pandata <- cbind(Y, x, ID)
  pandata <- pdata.frame(pandata, index = "ID")

  within <- plm(Y ~ x, data = pandata, model = "within")
  random <- plm(Y ~ x, data = pandata, model = "random")

  withCoef <- within$coefficients
  randCoef <- random$coefficients[2]
  
  return(randCoef)

}

data2 <- replicate(100, sim())

fullData <- cbind(data, data2)

sim <- function(){
  
  n = 100
  t = 100
  a = rep(rnorm(n), each = t)
  x <- (a + rnorm(n*t)) / 2
  u <- rnorm(100)
  B0 <- 2
  B1 <- 5

  Y <- B0 + B1 * x + a + u

  ID <- rep(1:n, each = t)
  pandata <- cbind(Y, x, ID)
  pandata <- pdata.frame(pandata, index = "ID")

  within <- plm(Y ~ x, data = pandata, model = "within")
  random <- plm(Y ~ x, data = pandata, model = "random")

  withCoef <- within$coefficients
  randCoef <- random$coefficients[2]
  
  test <- phtest(within, random)
  
  return(test)

}

test <- replicate(100, sim())

```

When the Covariance between X and A is not equal to zero, the coefficients between the two models vary widely. Furthermore, the standard error is not all that close to 0, and the Hausman test succeeds (p < .05). When N and T are increased to 100, the estimates still remain far apart, the standard errors are slightly big, and the Hausman test succeeds.

Part 2

1.

```{r}

data <- read.csv("https://raw.githubusercontent.com/PrisonRodeo/PLSC504-2023-git/main/Exercises/PLSC504-2023-ExerciseFive.csv")

pandata <- pdata.frame(data, index = c("Country", "Year"))
pandata <- drop_na(pandata)

#phtest(within, random)
within <- plm(TotalTrade ~ UnivEnrollmentPerK + POLITY + GDP + PercentLiterate, data = pandata, model = "within")

summary(within)

```

My model suggests that University Enrollment, Polity, GDP, and Literacy all play an important role in predicting a nation's total trade. For a one unit increase in University Enrollment, total trade decreases by -5660. For a one unit increase in polity, total trade decreases by -3459. For a one unit increase in GDP, total trade increases by 69 dollars. For a one unit increase in percent literate, total trade increases by 2056. 
