---
title: "PLSC 504 Hemauer Assignment #3"
output: html_notebook
---

Part 1

1.

```{r}

library(dplyr)
library(MatchIt)
library(marginaleffects)
library(car)

set.seed(1337)

repData <- 1
repData[1:10000] <- sample(0:1, 10000, replace = TRUE, prob = c(.95, .05)) #Treatment-- Phone Call (1 or 2) and control

repData2 <- repData
numChange <- round(.35 * length(repData))    #These lines correlate the treatment with the voting column
indicesChange <- sample(1:length(repData), numChange)
repData2 <- ifelse(seq_len(length(repData2)) %in% indicesChange, 0, repData2) #Vote

repData3 <- 0
repData3[1:10000] <- sample(0:1, 10000, replace = TRUE, prob = c(.98, .02)) #Newly registered voter

repData4 <- 0
repData4[1:10000] <- sample(18:90, 10000, replace = TRUE) #Age

repData5 <- 0
repData5[1:10000] <- sample(0:1, 10000, replace = TRUE, prob = c(.45, .55)) #Female

fullData <- as.data.frame(cbind(repData, repData2, repData3, repData4, repData5))
cor(fullData$repData, fullData$repData2)

model <- matchit(data = fullData, repData ~ repData3 + repData4 + repData5, method = "exact", distance = "glm")

sumModel <- match.data(model)

model3 <- lm(repData2 ~ repData * (repData3 + repData4 + repData5), data = sumModel, weights = weights)
avg_comparisons(model3, variables = "repData", vcov = ~subclass, newdata = subset(sumModel, repData == 1), wts = "weights")

```


Part 2

1.

```{r}

data <- read.csv("https://raw.githubusercontent.com/PrisonRodeo/PLSC504-2023-git/main/Exercises/PLSC504-2023-ExerciseThree.csv")

#murder ~ deathpen + urban + estbkpct + pop + aveeduc + aveideol

model1 <- lm(data = data, murder ~ deathpen + urban + estbkpct + pop + aveeduc + aveideol)
summary(model1)

```

This linear regression model finds that the death penalty is not predictive of murder rate. How urban a state is, the state's black population, and total population are the most predictive of murder rate... with each being a positive coefficient suggesting that when each is increased, murder increases. Average education and average ideology are less significant but still significant nonetheless. 

2.

```{r}

model2 <- matchit(data = data, deathpen ~ urban + estbkpct + pop + aveeduc + aveideol, method = "full", distance = "glm")

sumModel <- match.data(model2)

model3 <- lm(murder ~ deathpen * (urban + estbkpct + pop + aveeduc + aveideol), data = sumModel, weights = weights)
avg_comparisons(model3, variables = "deathpen", vcov = ~subclass, newdata = subset(sumModel, deathpen == 1), wts = "weights")

```

My results suggest that the deathpenalty has a -3.29 effect on the murder rate.

3.

Unsure how to test the model other than by checking the balance. It could also be tested by varying the method of matching used (as a sensitivity analysis).

4.

```{r}

data$lowcont[which(data$lowcont < 1)] <- 0

model2 <- matchit(data = data, lowcont ~ pop + aveeduc + urban, method = "full", distance = "glm")

sumModel <- match.data(model2)

model3 <- lm(murder ~ lowcont * (urban + pop + aveeduc), data = sumModel, weights = weights)
avg_comparisons(model3, variables = "lowcont", vcov = ~subclass, newdata = subset(sumModel, lowcont == 1), wts = "weights")

```

The first result shows that murder rate increases if the House is Democratic.

```{r}

data$upcont[which(data$upcont < 1)] <- 0

model2 <- matchit(data = data, upcont ~ pop + aveeduc + urban, method = "full", distance = "glm")
summary(model2)

sumModel <- match.data(model2)
head(sumModel)

model3 <- lm(murder ~ upcont * (urban + pop + aveeduc), data = sumModel, weights = weights)
summary(model3)
avg_comparisons(model3, variables = "upcont", vcov = ~subclass, newdata = subset(sumModel, upcont == 1), wts = "weights")

```

The second result shows that murder rate increases if the Senate is Democratic.